---
title: "36-315 Homework 01, Fall 2016"
author: "Solutions"
date: "Due Wednesday, January 25, 2016 (12pm ET) on Blackboard"
output: 
  html_document:
    toc:  true
    toc_float:  true
    code_folding:  show
---

**Style Guide**:  We are using Hadley Wickham's style guide.


Note that we include a floating table of contents in the homework solutions, to allow the viewer easier navigation between problems.  See the code in the header of the document for how to do this, and please feel free to include this in your future assignments!

#  Problem 1

**Customizing the RStudio User Interface:**

Personally, Sam prefers to have:

+  Source in its own pane on the top-right-hand side
+  Console in it's own pane on the top-left-hand side of the window
+  Help and Git ('VCS' for version control system) on the bottom-right
+  Everything else on the bottom-left

Sam personally prefers a fixed-width font like Courier.  Finally, when coding, Sam prefers to minimize the bottom-left and bottom-right panes so that he has more vertical space to see the code he's writing and the output it's generating.  Sam's also okay with swapping the Source and Console windows above, though this obviously is up to you.

Essentially any configuration is acceptable with justification. 


***
***


#  Problem 3

**Writing R Functions and Working with Vectors**:  

### a.

```{r}
quadratic <- function(a, b, c, x) {
  return(a * x^2 + b * x + c)
}
```

### b.

```{r}
# example tests
all(
  quadratic(1, 5, 2, 0) == 2,
  quadratic(0, 5, 2, 1) == 7,
  quadratic(1, 0, 1, 10) == 101,
  quadratic(2, 3, 0, 5) == 65,
  quadratic(-1, 5, 5, 5) == 5
)
# note:  the all() function returns TRUE if all input expressions are true, and FALSE otherwise
```

Our function works correctly on a variety of examples, as demonstrated above.

### c.

We get an error message: `Error in quadratic(x = 2) : argument "a" is missing, with no default Calls`.

### d.

```{r}
quadratic2 <- function(a=1, b=1, c=1, x) {
  return(a * x^2 + b * x + c)
}
quadratic2(x=2)
```
Since the other arguments were given default values, R uses the default values along with the one specified argument and runs successfully.

### e.   
```{r}
1:10
```
We get a vector with all of the integers between 1 and 10, inclusive.

### f. 
```{r}
quadratic2(x = 1:10)
```
We get a vector with the output of `quadratic2` for each of the values in `1:10` (and the given default values for the coefficients). This happens because R does element-wise arithmetic on vectors.

### g.  
```{r fig.width=5, fig.height=5, fig.align='center'}
plot(1:10, quadratic2(x = 1:10))
```
The graph is a scatterplot, with x values from the vector `1:10` and the output of `quadratic2` as the corresponding y values.  The graph does not have appropriate axis labeling or a descriptive title.


***
***

#  Problem 4

**Introduction to ggplot2**:  

### a.  
```{r}
#install.packages("ggplot2")
#library(ggplot2)

library(tidyverse)
```

NB: Do not include uncommented `install.packages` commands in your code.

### b.

"gg" stands for "Grammar of Graphics."

### c.  

The authors of `ggplot2` are Hadley Wickham and Winston Chang. Hadley wickham maintains `ggplot2`. Other `R` packages he has written include `dplyr`, `testthat`, `reshape2`, and 'devtools`. 


### d. 

```{r}
library(MASS)
data(Cars93)

head(Cars93)
ncol(Cars93)
colnames(Cars93)
```

### e. 

```{r, warning = FALSE, message = FALSE}
library(tidyverse)
library(MASS)
data(Cars93)

#  First graph
ggplot(data = Cars93) + 
  geom_bar(aes(x = Type), fill = "darkblue", color = "black") +
  labs(title = "Distribution of Car Types", 
       x = "Type of Car",
       y = "Number of Models",
       caption = "Source: R MASS Package",
       subtitle = "Cars93 Data Set") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))

#  Second graph
ggplot(data = Cars93) + 
  geom_histogram(aes(x = MPG.city), fill = "orange", color = "black") +
  labs(title = "City MPG vs. Fuel Tank Capacity",
       caption = "Cars93 Data Set",
       x = "Fuel Tank Capacity (gal.)",
       y = "City Miles per Gallon") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))

#  Third graph
ggplot(data = Cars93) + 
  geom_point(aes(x = Fuel.tank.capacity, y = MPG.highway, color = Type)) +
  labs(title = "Highway MPG vs. Fuel Tank Capacity",
       caption = "Cars93 Data Set",
       subtitle = "By Car Type",
       x = "Fuel Tank Capacity (gal.)",
       y = "Highway Miles per Gallon") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))

#  Fourth graph
ggplot(data = Cars93) + 
  geom_text(aes(x = Fuel.tank.capacity, y = MPG.city, 
                label = Type, color = Origin)) +
  labs(title = "City MPG vs. Fuel Tank Capacity",
       subtitle = "By Country of Origin and Car Type", 
       caption = "Cars93 Data Set",
       x = "Fuel Tank Capacity (gal.)",
       y = "City Miles per Gallon",
       color = "Country of Origin") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
```

Overall, the graph indicates a negative association between fuel tank capacity and city miles per gallon in the Cars93 data set; models with larger fuel tank capacities tend to have lower city miles per gallon and vice versa. This makes sense in theory, as one could imagine that larger fuel tank capacities are associated with larger cars, which get fewer miles per gallon. It also seems that American car models tend to have fewer city miles per gallon, as there seem to be more points correpsonding to U.S. cars in the lower part of the graph and more points corresponding to foreign cars in the upper part of the graph. With respect to car type, small cars tend to have smaller fuel tank capacities and larger city miles per gallon, as most of the small cars are on the top left of the graph. On the other hand, large cars and vans tend to have larger fuel tank capacities and lower city miles per gallon, as they are mainly on the bottom right of the graph.

***
***
