---
title: "36-315 Homework 01, Fall 2016"
author: "Solutions"
date: "Due Wednesday, January 25, 2016 (12pm ET) on Blackboard"
output: 
  html_document:
    toc:  true
    toc_float:  true
    code_folding:  show
---

**Style Guide**:  We are using Hadley Wickham's style guide.


Note that we include a floating table of contents in the homework solutions, to allow the viewer easier navigation between problems.  See the code in the header of the document for how to do this, and please feel free to include this in your future assignments!


##  Problem 1

Personally, Sam prefers to have:

+  Source in its own pane on the top-right-hand side
+  Console in it's own pane on the top-left-hand side of the window
+  Help and Git ('VSC' for version control system) on the bottom-right
+  Everything else on the bottom-left

Sam personally prefers a fixed-width font like Courier.  Finally, when coding, Sam prefers to minimize the bottom-left and bottom-right panes so that he has more vertical space to see the code he's writing and the output it's generating.  Sam's also okay with swapping the Source and Console windows above, though this obviously is up to you.

Essentially any configuration is acceptable with justification. 


***
***


##  Problem 3

**Writing R Functions and Working with Vectors**:  

a.  (5 points)  Write an R function called `quadratic` that computes the quadratic formula.  That is, write a function that has four inputs -- ```a```, ```b```, ```c```, and ```x``` and returns the quantity $ax^2 + bx + c$.

```{r}
quadratic <- function(a, b, c, x) {
  return(a * x^2 + b * x + c)
}
```

b.  (2 points)  Test your function out and demonstrate that it works for at least there different combinations of the parameters.

```{r}
# example tests
all(
  quadratic(1, 5, 2, 0) == 2,
  quadratic(0, 5, 2, 1) == 7,
  quadratic(1, 0, 1, 10) == 101,
  quadratic(2, 3, 0, 5) == 65,
  quadratic(-1, 5, 5, 5) == 5
)
# note:  the all() function returns TRUE if all input expressions are true, and FALSE otherwise
```

Our function works correctly on a variety of examples, as demonstrated above.

c.  (2 points)  Type `quadratic(x=2)` into your code block.  What happens when you only specify one argument?

We get an error message: `Error in quadratic(x = 2) : argument "a" is missing, with no default Calls`.

d.  (2 points)  Create a new function, `quadratic2`, that has default values for ```a=1```, ```b=1```, and ```c=1```.  Type `quadratic2(x=2)` into your code block.  What happens when you only specify one argument?

```{r}
quadratic2 <- function(a=1, b=1, c=1, x) {
  return(a * x^2 + b * x + c)
}
quadratic2(x=2)
```
Since the other arguments were given default values, R uses the default values along with the one specified argument and runs successfully.

e.  (2 points)  Type `1:10` at the command line in R.  What happens?  
```{r}
1:10
```
We get a vector with all of the integers between 1 and 10, inclusive.

f.  (2 points)  What happens when you call the function with the following input:  `quadratic2(x = 1:10)`.  Why does this happen?
```{r}
quadratic2(x = 1:10)
```
We get a vector with the output of `quadratic2` for each of the values in `1:10` (and the given default values for the coefficients). This happens because R does element-wise arithmetic on vectors.

g.  (5 points)  Type `plot(1:10, quadratic2(x=1:10))` into your code block.  Describe the graph that shows up.
```{r fig.width=5, fig.height=5, fig.align='center'}
plot(1:10, quadratic2(x = 1:10))
```
The graph is a scatterplot, with x values from the vector `1:10` and the output of `quadratic2` as the corresponding y values.  The graph does not have appropriate axis labeling or a descriptive title.


***
***
