---
title: "Lecture 21 R Demo -- Maps"
author: "Sam Ventura"
date: "April 5th, 2017"
output: 
  html_document:
    toc:  true
    toc_float:  true
    code_folding:  show
---


The graphs below don't have proper titles, axis labels, legends, etc.  Please take care to do this on your own graphs.

***


#  Adding to Static Maps

We can create static maps (road maps, satellite view maps, hybrid maps) with the `ggmap` package:

```{r}
library(tidyverse)
library(ggmap)
map_base <- get_map(location = c(lon = -79.944248, lat = 40.4415861),
                    color = "color",
                    source = "google",
                    maptype = "road",
                    zoom = 6)

map_object <- ggmap(map_base,
                    extent = "device",
                    ylab = "Latitude",
                    xlab = "Longitude")
map_object
```


Here, we'll work with airline data from [this GitHub repository](https://github.com/jpatokal/openflights).

Max Marchi wrote a great summary of using maps in R with ggmap [here](http://www.milanor.net/blog/maps-in-r-plotting-data-points-on-a-map/).  We'll follow this example closely in today's lecture.  Thanks, Max!

Before we begin, note that this is just one example of how you can add interesting information to maps with `ggmap`.  As long as you have latitude and longitude information, you should be able to add data to maps.  For more interesting examples and for an in-depth description of `ggmap`, see the short paper by David Kahle and Hadley Wickham [here](http://stat405.had.co.nz/ggmap.pdf).


##  Load flight data from GitHub

To load data from GitHub, you should navigate to the raw file, copy the URL, and use `read.csv()`.

```{r}
#  Load and format airports data
airports <- read_csv("https://raw.githubusercontent.com/jpatokal/openflights/master/data/airports.dat",
                     col_names = c("ID", "name", "city", "country", "IATA_FAA", 
                                   "ICAO", "lat", "lon", "altitude", "timezone", "DST"))
airports

#  Load and format routes data
routes <- read_csv("https://raw.githubusercontent.com/jpatokal/openflights/master/data/routes.dat",
                   col_names = c("airline", "airlineID", "sourceAirport", 
                                 "sourceAirportID", "destinationAirport", 
                                 "destinationAirportID", "codeshare", "stops",
                                 "equipment"))
routes
```


##  Manipulating the data to get some custom data.frames

Here, we'll do some data manipulation to obtain the number of arrivals/departures per airport.

```{r}
#  Manipulate the routes data to create two new data.frames
#    one for arrivals, one for departures.  
#  Each counts the number of flights arriving/departing from each airport.
departures <- routes %>%
  group_by(sourceAirportID) %>%
  summarize(flights = n()) %>%
  mutate(sourceAirportID = as.integer(as.vector(sourceAirportID)))

arrivals <- routes %>%
  group_by(destinationAirportID) %>%
  summarize(flights = n()) %>%
  mutate(destinationAirportID = as.integer(as.vector(destinationAirportID)))

#  Merge each of the arrivals/departures data.frames with the airports data.frame above
airportD <- left_join(airports, departures, by = c("ID" = "sourceAirportID"))
airportA <- left_join(airports, arrivals, by = c("ID" = "destinationAirportID"))

#  Create data.frame of routes to/from a specific city
my_airport_code <- "PIT"
my_routes <- filter(routes, 
                    sourceAirport == my_airport_code | 
                      destinationAirport == my_airport_code)

#  Add in relevant information from the airports data.frame
#  Do this in two steps, so that you first pull in the source airport information
#  and then pull in the destination airport information
#  This can be done easily with two calls to left_join()
#  The rest of the code is just formatting
my_airport_data <- my_routes %>%
  left_join(airports, by = c("sourceAirport" = "IATA_FAA")) %>%
  select(destinationAirport, lat, lon, timezone) %>%
  rename(source_lat = lat, source_lon = lon, source_timezone = timezone) %>%
  left_join(airports, by = c("destinationAirport" = "IATA_FAA")) %>%
  select(source_lat, source_lon, source_timezone, lat, lon, timezone) %>%
  rename(dest_lat = lat, dest_lon = lon, dest_timezone = timezone)
```


##  Mapping the data we created

We'll use `ggmap` to create a base map of the data.  Now, we'll add layers of points onto the map.

```{r}
#  Use ggmap to visualize the European flights
library(ggmap)
map <- get_map(location = 'Europe', zoom = 4)
map <- get_map(location = 'United States', zoom = 4)

#  Visualize the basic map
ggmap(map)

#  Add points to the map of departures
#  Each point will be located at the lat/long of the airport
#  The size of the points is proportional to the square root of the number of flights at that airport
mapPoints <- ggmap(map) +
  geom_point(aes(x = lon, y = lat, size = sqrt(flights), color = DST), 
             data = airportD, alpha = .5)

#  Add a custom legend to the plot
mapPointsLegend <- mapPoints +
  scale_size_area(breaks = sqrt(c(1, 5, 10, 50, 100, 500)), 
                  labels = c(1, 5, 10, 50, 100, 500), 
                  name = "departing routes")
mapPointsLegend
```


##  Further Data Manipulation and Facetting

Below, we create a new variable (flight type -- arrival or departure) and use it to create a single data.frame of all arrivals and departures.

We'll then facet on this variable so we can simultaneously visualize both arriving and departing flights.

```{r}
#  Create a data.frame containing both departures and arrivals
#  Do this by creating a "type" variable and then combining the two data.frames
#  We will later be able to use the type variable for facetting
airportD$type <- "departures"
airportA$type <- "arrivals"
airportDA <- bind_rows(airportD, airportA)

#  Create our base plot and add the custom legend
mapPointsDA <- ggmap(map) +
  geom_point(aes(x = lon, y = lat, size = sqrt(flights)), 
             data = airportDA, alpha = .5) +
  scale_size_area(breaks = sqrt(c(1, 5, 10, 50, 100, 500)), 
                  labels = c(1, 5, 10, 50, 100, 500), 
                  name = "routes")

#  Facet on flight type (arrival/departure)
mapPointsDA + facet_grid(. ~ type)
```



***
***


#  Geocoding with `ggmap`

```{r}
gc <- geocode("carnegie mellon university")
map <- get_map(gc, zoom = 6)
ggmap(map) + 
  geom_point(aes(x = lon, y = lat), 
             data = gc, size = 5, color = "red")
```


***
***


#  Using Spatial Polygons

What is a polygon?


##  Plotting spatial objects with `geom_polygon()`

In `ggplot()`, polygons are just another geometry, making it really easy to add geographic shapes (e.g. corresponding to countries, states, counties, etc.) to maps.

```{r}
#  Note:  The sp package can be really fussy at installation
#  If prompted, do not restart R when installing the package
#  If prompted with "Do you want to install from sources the packages which need compilation?", type 'n'
#install.packages("sp", dependencies = T)
library(sp)
library(ggmap)
us_data <- map_data("state")
county_data <- map_data("county")

us_county_map <- ggplot() + 
  geom_polygon(aes(long, lat, group=group), fill = "blue", size = 4, 
               data = county_data) + 
  geom_polygon(aes(long, lat, group=group), color = 'white',
               fill = NA, data = us_data) + 
  theme_bw() + theme(axis.text = element_blank(), 
                     axis.title = element_blank())
us_county_map
```


#  Map Projections

##  Use `coord_map()` to specify your map projection

Way back at the beginning of the semester, we learned about using `coord_cartesian()` and `coord_polar()` to specify the coordinates of our plots.  We're (finally) revisiting this idea in the context of geographic mapping.

Using the `coord_map()` function, we can specify what kind of map projection we want to use.

```{r}
us_county_map + coord_map("mercator") 
us_county_map + coord_map("polyconic") 
us_county_map + coord_map("sinusoidal") 
```


