---
title: "Maps Helper"
author:  "Jennifer Jin and Sam Ventura"
output: 
  html_document:
    toc:  true
    toc_float:  true
    code_folding:  show
---


#  Load the Data

```{r, warning = FALSE, message = FALSE}
library(tidyverse)
library(sp)
library(ggmap)
museums <- read_csv("https://raw.githubusercontent.com/sventura/315-code-and-datasets/master/data/museums.csv")
```


#  Choropleth Maps

Create a choropleth map of the United States that shows the number of musuems in each state.

```{r, warning = FALSE, message = FALSE}
us_data <- map_data("state")

museums_by_state <- museums %>% 
  group_by(State) %>%
  summarize(Count = n(),
            log_mean_income = log(mean(Income, na.rm = T)))

us_data <- us_data %>%
  left_join(museums_by_state, by = c("region" = "State"))

ggplot(us_data) +
  geom_polygon(aes(x = long, y = lat, group = group, fill = Count), 
               color = "black") + 
  scale_fill_continuous(low = "green", high = "purple") +
  theme_void() +
  coord_map("polyconic") + 
  labs(title = "Number of Museums Per State")

ggplot(us_data) +
  geom_polygon(aes(x = long, y = lat, group = group, 
                   fill = log_mean_income), color = "black") + 
  scale_fill_continuous(low = "green", high = "purple") +
  theme_void() +
  coord_map("polyconic") + 
  labs(title = "Log Mean Income of Museums Per State")
```


#  Layering (Colored) Points

Use `ggmap` to create a map centered on the continental United States. Add points corresponding to the location of each museum, colored by the type of museum.

```{r, warning = FALSE, message = FALSE}
library(ggmap)
map <- get_map(location = 'United States', zoom = 4)

#  Visualize the basic map
ggmap(map) +
  geom_point(aes(x = Longitude, y = Latitude, color = Type), 
             data = museums, alpha = .5) + 
  labs(title = "Museum Locations by Type") + 
  theme_void()
```

#  Filtering

Use the same base map in the previous section to create a map centered on the continental United States. Add points corresponding to natural history museums, sized by their income.

```{r, warning = FALSE, message = FALSE}
nat_hist_museums <- filter(museums, 
                           Type == "NATURAL HISTORY MUSEUM",
                           Income != 0)

ggmap(map) +
  geom_point(aes(x = Longitude, y = Latitude, size = Income), 
             data = nat_hist_museums, alpha = .5) + 
  labs(title = "Natural History Museums by Income") + 
  theme_void()
```





