---
title: "36-315 Lab 01"
author: "Sample Solution"
date: "Due Saturday, January 21st, 2017 (6:30pm)"
output: html_document
---


##  Lab 01:  Hello, world!  Introduction to R, RStudio, & RMarkdown


***
***


#  Problem 1

```{r}
#  I am printing ‘Hello, World!’ in R Markdown.
#  My major is *****
print("Hello, World! My name is *****.")
```


***
***


#  Problem 2

```{r}
#  a.  $(6+3)*4 - 5$
(6+3)*4 - 5

#  b.  $4^2$
4^2

#  c.  $e^{-5}$
exp(-5)

#  d.  Let $y=8$.  Let $x = 5-6y$.  Print $x$.
y = 8
x = 5-6*y
print(x)

#  e.  Repeat (d) for $y=0$, $y=1$, and $y=5/6$.
y = 0
x = 5-6*y
print(x)

y = 1
x = 5-6*y
print(x)

y=5/6
x = 5-6*y
print(x)
```


***
***


#  Problem 3

a.  
```{r}
help.search("quantile")
```

numeric vector of probabilities with values in [0,1]. (Values up to 2e-14 outside that range are accepted and moved to the nearby endpoint.)

b.  
```{r}
help.search("mean")
x <- c(0:10, 50)
xm <- mean(x)
c(xm, mean(x, trim = 0.10))
```

c.  
```{r}
help.search("standard deviation")
sd(x, na.rm = FALSE)
```


***
***


#  Problem 4

a.  
```{r}
library(MASS)
```

official name is "Modern Applied Statistics with S"

b.  
```{r}
library(datasets)
```

trees data set: Girth, Height and Volume for Black Cherry Trees

c.  
```{r}
library(graphics)
```

The function is called stripchart, x represents the data from which the plots are to be produced. In the default method the data can be specified as a single numeric vector, or as list of numeric vectors, each corresponding to a component plot. In the formula method, a symbolic specification of the form y ~ g can be given, indicating the observations in the vector y are to be grouped according to the levels of the factor g. NAs are allowed in the data.


***
***


#  Problem 5

A lot of install messages were printed out in the console.

At the end the console printed out the installed package location in my computer.


***
***


#  Problem 6

```{r}
#  Load the tidyverse library
library(tidyverse)

#  Read in the data
bridges <- read_csv("https://raw.githubusercontent.com/sventura/315-code-and-datasets/master/data/bridges-pgh.csv")

#  Create the new variables
bridges <- mutate(bridges, 
                  over_allegheny = ifelse(river == "A", "yes", "no"),
                  length_binary = ifelse(length >= 3000, "long", "short"))

#  Create the bar plot
ggplot(bridges, aes(x = river)) + 
  geom_bar(fill = "darkblue") + 
  labs(
    title = "Number of Bridges in Pittsburgh by River",
    subtitle = "A = Allegheny River, M = Monongahela River, O = Ohio River",
    x = "River",
    y = "Number of Bridges",
    caption = "Source:  Pittsburgh Bridges Dataset"
  )
```


***
***

