---
title: "36-315 Homework 03, Fall 2016"
author: "Solutions"
date: "Due Wednesday, September 21, 2016 (12pm ET) on Blackboard"
output: 
  html_document:
    toc:  true
    toc_float:  true
    code_folding:  show
---

##  Homework 03:  Data Manipulation for 2+ Dimensional Categorical Graphics


#  Problem 1


a.  The author discusses the data, the geometries, the scaling and the coordinates (polar vs. Cartesian).

A possible transformation to the x- or y-axis scaling in `ggplot()` is a log transformation.  The way to apply a logarithmic scale is adding `scale_x_log10()`. 

The author asserts that `ggplot()` is better for creating facetted graphs because in base R you have to write out separate commands for each split by creating an entirely separate dataset for each partition.  Also in `ggplot()`, facetting keeps identifcal axis limits for each partition and allows for easy aesthetic changes. 


b.  The geometry used to create a scatterplot of two continuous variables is geom_point(). 

c.  The author lists four "advanced parts" of the `ggplot2` grammar

  + Statistical transformations: plots statistical summaries of the data, like quantiles, fitted curves, sums 
  + Coordinate Systems: mapping data coordinates into the plane of the data rectangle
  + Facets: Arrangement of the data into a grid of plots 
  + Visual Themes: Overall visual defaults of a plot: background, grids, axe, default typeface, sizes, colors etc.
  

***
***



#  Problem 2

**Critiquing Graphs --- xkcd**:

Take a look at [this graphic](http://xkcd.com/1732/) from xkcd.  

a.  The graph shows the change in the global average temperature over the past ~22000 years. Years are plotted on the vertical axis, while the horizontal axis shows the difference between the global average temperature of the time on the vertical axis and the global average temperature between 1961 and 1990. The background color gradient also represents the difference between the global average temperature and the global average temperature between 1961 and 1990. 

b.  The graph does a decent, though not perfect job, of achieving its goals. The line on the graph clearly shows that, while global average temperature has fluctuated throughout history, it has not done so with the kind of magnitude and speed that we have experienced since the beginning of the 21st century. The change in the temperature corresponding to the line is fairly easy to read and interpret, and there is no confusion about which variables are being plotted or how they are represented. The graph also does not distort the data. The graph has some unnecessary data ink. There are some artistic drawings which do not need to be there and distract the reader,from the point of the plot. The background color gradient is redundant with the horizontal axis. It is hard to look at the graph in its entirety due to its size, since one needs to scroll a long way to see the entire thing. If I could change this graphic, I would get rid of the background color and the drawings in the middle, and perhaps shrink the vertical axis to make the plot more viewable. 

c.  Most of the text does not help increase the reader's understanding of the graphic, and, if anything, only serve to distract the reader. As mentioned previously, the drawings are certainly unnecessary. Many of the text annotations include historical events that do not have anything to do with climate change, and thus do not serve to inform the reader about why we see the trends in the plot. There is also a redundancy in the y-axis labeling, telling us which direction is colder and which is warmer. The caption does provide a decent narrative about what the plot is attempting to show and the point that it is tryin to make. I would include in the surrounding text what the data actually is, as we do not know how long of a time period that each point on the line is averaged over. 

d.  The main issue with the graphic that the NPR article brings up is that there are actually two different data sets being displayed; one is the measured yearly average, which has been taken over the past 150 years, while the other is a reconstruction of a running average of Earth's temperature over a scale of a few hundred years. Thus, large but quick spikes would be smoothed out of the top of the graph. 

e.  xkcd isn't funny.  End of story. 


***
***


#  Problem 3

**More on Pie Charts and Rose Diagrams**

a.  

```{r, warning = F, message = F}
library(data.table)
library(ggplot2)
library(gridExtra)
airline_tweets <- fread("https://raw.githubusercontent.com/sventura/315-code-and-datasets/master/data/Tweets.csv")

ggplot(airline_tweets) +
  aes(x = factor(1), fill = airline) +
  geom_bar(width = 1) + 
  coord_polar(theta = "y") +
  ggtitle("Tweet counts by airline") + 
  xlab("") + ylab("") + labs(fill = "Airline")
  scale_x_discrete(breaks = NULL)
## width = 1 in geom_bar() removes the whole in the center
## scale_x_discrete() removes the tick mark for the x variable

```

b.  
```{r, warning = F, message = F, fig.width= 12, fig.height= 6}
p.stackedbar <- 
  ggplot(data = airline_tweets, aes(x = airline, fill = airline_sentiment)) +
  geom_bar() +  ggtitle("Tweet Counts by Airline") +
  labs(x = "", y = "Number of Tweets") + 
  scale_fill_discrete(guide = guide_legend(title = "Tweet Sentiment"))
p.stackedrose <- p.stackedbar + coord_polar() 
grid.arrange(p.stackedbar  +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)),
  p.stackedrose + labs(x = "", y = ""),
  nrow = 1)
```


c.  Describe and critique the stacked rose diagram.  

+  The marginal distribution of `airline` is shown by the radius of each segment.

+  The conditional distribution of `airline_sentiment` is given that the tweet was about Delta is shown in the piece corresponding to Delta.  Where the conditional distribution is the proportion of the total radius of each slice (not the area) made up by tweets with a given sentiment.

+  First, as is generally the case with a rose diagram area does not correspond to total. If, for example, we were to switch the order of the labels so that tweets with negative were displayed on the outside the result would be that a much greater area in the graph was devoted to these tweets even though the data would have stayed the same.  Second, polar coordinates make it hard to compare the relative totals of tweets about airlines that do not happen to be plotted next to eachother.

***
***



#  Problem 4

(3 points each)

**Bar Charts at the Proportion and Percentage Scales**


```{r, warning = F, message = F}
library(data.table)
library(ggplot2)
bike_data <- fread("https://raw.githubusercontent.com/sventura/315-code-and-datasets/master/data/nyc-citi-bike-data-subset.csv")
dim(bike_data)
```


a.  The graphs are the same.

```{r}
ggplot(data = bike_data, aes(x = gender)) + 
  geom_bar(aes(y = ..count.., fill = gender)) + 
  labs(x = "Gender", y = "Number of Riders") + 
  ggtitle("Gender of NYC Citi Bike Riders") + 
  scale_fill_discrete(name = "Gender", labels = c("Unknown", "Male", "Female"))
```

b.  

```{r}
ggplot(data = bike_data, aes(x = gender)) + 
  geom_bar(aes(y = (..count..)/sum(..count..), fill = gender)) + 
  labs(x = "Gender", y = "Proportion of Riders") + 
  ggtitle("Gender of NYC Citi Bike Riders") + 
  scale_fill_discrete(name = "Gender", labels = c("Unknown", "Male", "Female"))

```

c.  

```{r}
ggplot(data = bike_data, aes(x = gender)) + 
  geom_bar(aes(y = 100 * (..count..)/sum(..count..), fill = gender)) +
  labs(x = "Gender", y = "Percentage of Riders") + 
  ggtitle("Gender of NYC Citi Bike Riders") + 
  scale_fill_discrete(name = "Gender", labels = c("Unknown", "Male", "Female"))
```

d.  
```{r}
library(scales)

ggplot(data = bike_data, aes(x = gender)) + 
  geom_bar(aes(y = (..count..)/sum(..count..), fill = gender)) +
  scale_y_continuous(labels = percent) + 
  labs(x = "Gender", y = "Proportion of Riders") + 
  ggtitle("Gender of NYC Citi Bike Riders") + 
  scale_fill_discrete(name = "Gender", labels = c("Unknown", "Male", "Female"))

```


***
***




#  Problem 5

(12 points)

**Data Manipulation and Reordering the Bars**

```{r, warning = F, message = F, fig.width = 15}
library(data.table)
library(dplyr)
library(ggplot2)
library(gridExtra)
airline_tweets <- fread("https://raw.githubusercontent.com/sventura/315-code-and-datasets/master/data/Tweets.csv")

#  Manipulate the airline 
airline_counts <- airline_tweets %>%  #  Start with the airline_tweets data.frame
  group_by(airline) %>%  #  group by the airline variable
  summarise(count = n()) %>%  #  summarize the dataset by calculating the count of each airline
  mutate(total = sum(count))  #  add a column with the total number of flights to the result

#  The result is a new dataset that has three columns:  airline, count, and total
print(airline_counts)

#  Basic bar graph with ggplot
airline_tweet_plot <- ggplot(data = airline_counts, aes(x = airline, y = count)) + 
  geom_bar(stat="identity", fill = "#008080", color = "#FF8C00") +
  ggtitle("Number of Tweets by Airline") +
  labs(x = "Airline", y = "Number of Tweets")

#  Reorder the bars in increasing order:
airline_tweet_increasing_plot <- ggplot(data = airline_counts, aes(x = reorder(airline, count), y = count)) + 
  geom_bar(stat="identity", fill = "#008080", color = "#FF8C00") +
  ggtitle("Number of Tweets by Airline") +
  labs(x = "Airline", y = "Number of Tweets")

#  Reorder the bars in increasing order:
#  Note: We will learn more about piping [%>%] throughout the course. 
#  There are multiple ways to accomplish this
airline_tweet_decreasing_plot <- ggplot(data = airline_counts, aes(x = reorder(airline, count %>% (function(x) return(-x))), y = count)) + 
  geom_bar(stat="identity", fill = "#008080", color = "#FF8C00") +
  ggtitle("Number of Tweets by Airline") +
  labs(x = "Airline", y = "Number of Tweets")

#  Plot all 3
grid.arrange(airline_tweet_plot, airline_tweet_increasing_plot, airline_tweet_decreasing_plot, nrow = 1)
```


***
***


#  Problem 6

(4 points each)

**More Data Manipulation and Incorporating Statistical Information Into Graphs**

Using the same set-up as Problem 5, we'll add some additional information to `airline_counts`:

```{r}
#  Add the following information to the airline_counts dataset:
#    Proportions and percentages corresponding to each category
#    The standard error on the proportios or percentages corresponding to each category
#    Lower bound of an (approximate) 95% confidence interval around the true proportion in each category
#    Upper bound of an (approximate) 95% confidence interval around the true proportion in each category
airline_counts <- mutate(airline_counts,
                         proportion = count / total,
                         percentage = proportion * 100,
                         std_error = 87,
                         lower = proportion - 1.96 * std_error,
                         upper = proportion + 1.96 * std_error)

print(airline_counts)
```

a.

```{r}
airline_counts <- mutate(airline_counts,
                         proportion = count / total,
                         percentage = proportion * 100,
                         std_error =  sqrt(proportion * (1 - proportion) 
                                           / total),
                         lower = proportion - 1.96 * std_error,
                         upper = proportion + 1.96 * std_error)

print(airline_counts)
```

b.

```{r fig.width = 10}
#  Bar graph of percentages
airline_percentage_plot <- ggplot(data = airline_counts, aes(x = airline, y = percentage)) + 
  geom_bar(stat="identity", fill = "#008080", color = "#FF8C00") +
  ggtitle("Percentage of Total Tweets per Airline") +
  labs(x = "Airline", y = "Percentage of Tweets")

#  Bar graph of proportions
airline_proportion_plot <- ggplot(data = airline_counts, aes(x = airline, y = proportion)) + 
  geom_bar(stat="identity", fill = "#008080", color = "#FF8C00") +
  ggtitle("Proportion of Total Tweets per Airline") +
  labs(x = "Airline", y = "Proportion of Tweets")

grid.arrange(airline_percentage_plot, airline_proportion_plot, nrow = 1)
```

The difference between these two graphs is the variable plotted on the y-axis of the plot, which changes its scale. The one on the left shows the percentage of tweets that each airline was mentioned in, out of 100, while the one on the right shows the proportion of tweets that each airline was mentioned in, out of 1. 

c.

```{r}
airline_proportion_error_plot <- ggplot(data = airline_counts, aes(x = airline, y = proportion)) + 
  geom_bar(stat="identity", fill = "#008080", color = "#FF8C00") +
  ggtitle("Proportion of Total Tweets per Airline") +
  labs(x = "Airline", y = "Proportion of Tweets") +
  geom_errorbar(aes(ymin = lower, ymax = upper))

airline_proportion_error_plot
```

The 95% confidence intervals of the proportion of tweets about US Airways and American Airlines intersect (as can be confirmed by looking at the data frame.) Thus, there is not a statistically significant difference between the proportions of tweets about American and US Airways.

***
***


#  Problem 7

(3 points each)

**Even More Data Manipulation with the NYC Citi Bike Data**


```{r, message = F, warning = F}
library(reshape2)
date_time <- colsplit(bike_data$starttime, " ", c("date", "time"))
bike_data <- mutate(bike_data, 
                    birth_decade = paste0(substr(`birth year`, 1, 3), "0s"),
                    date = as.vector(date_time$date),
                    time_of_day = as.vector(date_time$time),
                    hour_of_day = substr(time_of_day, 1, 2),
                    am_or_pm = ifelse(as.integer(hour_of_day) < 12, "AM", "PM"),
                    day_of_week = weekdays(as.Date(date, format = "%m/%d/%Y")),
                    less_than_30_mins = ifelse(tripduration < 1800,
                                               "Short Trip", "Long Trip"),
                    weekend = ifelse(day_of_week %in% c("Saturday", "Sunday"),
                                     "Weekend", "Weekday"))
```


a.

```{r, message = F, warning = F}
ggplot(bike_data, aes(x = day_of_week)) +
  geom_bar() +
  labs(x = "Day of the Week", y = "Number of Rides",
       title = "Number of Bikeshare rides by Day")
```

The days are not displayed in the order we would expect

b.

```{r, message = F, warning = F}
bike_data$day_of_week <- 
  factor(bike_data$day_of_week,
         levels= c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday"))

ggplot(bike_data, aes(x = day_of_week)) +
  geom_bar(fill = "red", color = "black") +
  labs(x = "Day of the Week", y = "Number of Rides",
       title = "Number of Bikeshare rides by Day")
```

There are generally more rides per weekend day than there are per week day.  There are some differences for specific weekdays as well.  For example, Wednesdays and Thursdays appear to have more frequent ridership than Mondays, Tuesdays, and Fridays.  However, upon further inspection, this is because the month of June began with a Wednesday and ended with a Thursday, meaning that there were five Wednesdays and Thursdays in June, while there were only 4 Mondays, Wednesdays, and Fridays.


c.

```{r, message = F, warning = F}
bike_data_subset <- filter(bike_data,
                        usertype != "Customer",
                        `birth year` >= 1950)

ggplot(bike_data_subset, aes(x = day_of_week)) +
  geom_bar(aes(fill = birth_decade)) +
  labs(x = "Day of the Week", y = "Number of Rides", fill = "Birth Decade",
       title = "Number of Bikeshare rides taken by Subscribers")
```

d.  We note that it is no longer the case that there are more rides taken on weekend days.  This suggests that the behavior of Customers and Subscribers to the bike share service follow different behavioral patterns.

e.  

```{r, message = F, warning = F}
bike_data_subset <- filter(bike_data,
                        usertype != "Customer",
                        `birth year` >= 1950)

ggplot(bike_data_subset, aes(x = day_of_week)) +
  geom_bar(aes(fill = birth_decade)) +
  labs(x = "Day of the Week", y = "Number of Rides",
       title = "Number of Bikeshare rides taken by Subscribers") +
  facet_grid(less_than_30_mins~am_or_pm) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  scale_fill_discrete(guide = guide_legend(title = "Birth Decade"))
```

Note that we use different scales for Long Trips vs. Short Trips as this makes is easier to compare distributions across groups.  People born in the 1990s also seem to account for a a greater share of both short and long trips made in the evening.  It also apperas that people born in the 1970s take an outsize share of short trips on Wednesday and Thursday mornings.

f. 
```{r, message = F, warning = F}
bike_data_subset <- filter(bike_data,
                        usertype != "Customer",
                        `birth year` >= 1950)

ggplot(bike_data_subset, aes(x = day_of_week)) +
  geom_bar(aes(fill = birth_decade)) +
  labs(x = "Day of the Week", y = "Number of Rides",
       title = "Number of Bikeshare rides taken by Subscribers") +
  facet_grid(less_than_30_mins~am_or_pm, scale = "free_y") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  scale_fill_discrete(guide = guide_legend(title = "Birth Decade"))
```

Differences we noted previously are now more apparent than they were previously.  Unfortunately, it is now substantially more difficult to compare the number of observations in bars across the facets (or visually estimate the total number of observations within a facet).

g.  (We did this in the previous part -- see the code above!)

h.

```{r, message = F, warning = F}
bike_data_subset <- filter(bike_data,
                        usertype != "Customer",
                        `birth year` >= 1950)

ggplot(bike_data_subset, aes(x = day_of_week)) +
  geom_bar(aes(fill = birth_decade), position="dodge") +
  labs(x = "Day of the Week", y = "Number of Rides",
       title = "Number of Bikeshare rides taken by Subscribers") +
  facet_grid(less_than_30_mins~am_or_pm, scale = "free_y") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  scale_fill_discrete(guide = guide_legend(title = "Birth Decade"))
```
 
The y-limits get smaller because they now only need to be large enough to show the largest single category.


***
***


#  Problem 8

(3 points each)

```{r, warning = F, message = F}
library(data.table)
library(dplyr)
library(reshape2)
tweets_sub <- filter(airline_tweets, 
                     user_timezone %in% c("Mountain Time (US & Canada)", 
                                          "Pacific Time (US & Canada)", 
                                          "Central Time (US & Canada)", 
                                          "Eastern Time (US & Canada)"))

temp <- colsplit(as.vector(tweets_sub$user_timezone), " ", c("first_word", "extra_word"))
tweets_sub <- mutate(tweets_sub, user_timezone = temp$first_word)
```

a.

```{r, warning = F, message = F}
tweets_sub <- mutate(tweets_sub,
                     user_timezone_we = factor(user_timezone,
                                               levels = c("Pacific", "Mountain", "Central", "Eastern")),
                     user_timezone_ew = factor(user_timezone,
                                               levels = rev(levels(user_timezone_we))))

p.west.east <- ggplot(tweets_sub, aes(x = airline)) +
  geom_bar(aes(fill = airline_sentiment), position="dodge") +
  facet_grid(.~user_timezone_we) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  scale_fill_discrete(guide = guide_legend(title = "Tweet Sentiment")) +
  labs(x = "Airline", y = "Number of Tweets",
       title = "Airline Tweets by Timezone West to East")

p.east.west <- ggplot(tweets_sub, aes(x = airline)) +
  geom_bar(aes(fill = airline_sentiment), position="dodge") +
  facet_grid(.~user_timezone_ew) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  scale_fill_discrete(guide = guide_legend(title = "Tweet Sentiment")) +
  labs(x = "Airline", y = "Number of Tweets",
       title = "Airline Tweets by Timezone East to West")

grid.arrange(p.west.east, p.east.west, ncol = 1)
```

I prefer the graph that goes from west to east as it is consistent with the way maps of the US are typically displayed.  That said, one could make an argument for the east-to-west graph, since this corresponds to the direction the sun moves (that is, mornings occur in the east first since the west is three hours behind).


b.  

```{r}
ggplot(tweets_sub, aes(x = airline)) +
  geom_bar(aes(fill = airline_sentiment), position="dodge") +
  facet_grid(user_timezone_we ~ ., margins = TRUE, scales = "free_y", as.table = FALSE) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  scale_fill_discrete(guide = guide_legend(title = "Tweet Sentiment")) +
  labs(x = "Airline", y = "Number of Tweets",
       title = "Airline Tweets by Timezone West to East")
```

It is easier to compare the marginal distributions if we switch to a column of graphs allowing the scales to adjust more.  In many cases the marginal distributions do not closely match the overall distribution.  For example US Airways has a much greater share in the Eastern timezone, and an offsetting smaller share in the Pacific and Mountain timezones than it does overall.  We can see a similar pattern for Delta.  In contrast American has a larger share of the tweets in the Central timezone than it does in the other timezones or the overall distribution.
