---
title: "Lecture 11 + R Demo"
author: "Sam Ventura"
date: "October 5th, 2016"
output: 
  html_document:
    toc:  true
    toc_float:  true
    code_folding:  show
---


#  Scatterplots!

So far, we've used the following data types within a single graphic:

+  1-D Categorical
+  2-D Categorical
+  3+D Categorical (with facets)
+  1-D Continuous
+  Mixing Categorical and 1-D Continuous


Now, we'll be looking at 2-D continuous, 3+D Continuous, and other high-dimensional data scenarios.

***

**The graphs below don't have proper titles, axis labels, legends, etc.  Please take care to do this on your own graphs.**

***

#  Geometry

We use `geom_point()` to generate scatterplots.

```{r}
library(MASS)
library(ggplot2)
data(Cars93)
#head(Cars93)

ggplot(Cars93) + aes(x = Fuel.tank.capacity, y = Price) + 
  geom_point()

```


#  Aesthetics

##  Colors

We can color by a third variable, whether it's categorical (different color for each category):

```{r}
ggplot(Cars93) + aes(x = Fuel.tank.capacity, y = Price) + 
  geom_point(aes(colour = Origin))
```

...or continuous (with a color scale/gradient):

```{r}
ggplot(Cars93) + aes(x = Fuel.tank.capacity, y = Price) + 
  geom_point(aes(colour = Weight))
```

##  Point size (`size`)

We can also map variables to other aesthetics, e.g. `size`:

```{r}
ggplot(Cars93) + aes(x = Fuel.tank.capacity, y = Price) + 
  geom_point(aes(size = Weight))
```


##  Point type (`shape`)

... or the type (`shape`) of points:

```{r}
ggplot(Cars93) + aes(x = Fuel.tank.capacity, y = Price) + 
  geom_point(aes(shape = Origin), size = 2)
```


##  Point shading (`alpha`)

... or the shading (`alpha`) of the points:

```{r}
ggplot(Cars93) + aes(x = Fuel.tank.capacity, y = Price) + 
  geom_point(aes(alpha = Weight), color = "red")
```


##  Combining aesthetics

We can even do several of these at once:

```{r}
ggplot(Cars93) + aes(x = Fuel.tank.capacity, y = Price) + 
  geom_point(aes(size = Weight, color = Type, shape = Origin,
                 alpha = Horsepower))
```

The above graph is actually a 6-dimensional graphic:

+  `x` = Fuel tank capacity
+  `y` = Price
+  `color` = Car type
+  `size` = Weight
+  `shape` = Origin
+  `alpha` = Horsepower



##  Mixing point types

Mixing point types is an easy way to make (aesthetically) nice looking graphics!


```{r}
ggplot(Cars93) + aes(x = Fuel.tank.capacity, y = Price) + 
  geom_point(colour="black", size = 4.5) +
  geom_point(colour="pink", size = 4) +
  geom_point(aes(shape = Origin))
```




###  But be careful!  

The more complications you add, the more difficult your graph is to explain.

***


#  Trend Lines

We can do different forms of statistical modeling **within our plots** (e.g. with `geom_smooth()`).

For example, we can add a linear regression line to our plot very easily.  

##  Linear regression

To do this, we can use `+ geom_smooth(method = lm)`:

```{r}
ggplot(Cars93) + aes(x = Fuel.tank.capacity, y = Price) + 
  geom_point() + geom_smooth(method = lm, se = FALSE)
```

##  Linear regression (with error bars)

Above, I added `se = FALSE` so that the standard error bars do not show up in the graph.  Setting this parameter to `TRUE` instead, we get:

```{r}
ggplot(Cars93) + aes(x = Fuel.tank.capacity, y = Price) + 
  geom_point() + geom_smooth(method = lm, se = TRUE)
```

This isn't a very good fit (e.g. look at residuals vs. fitted), so perhaps a variable transformation will help.


##  Variable Transformations {.tabset}

###  Log Transform

We can transform continuous variables as well -- again, **within the plot**.

```{r}
ggplot(Cars93) + aes(x = Fuel.tank.capacity, y = log(Price)) + 
  geom_point(aes(alpha = Weight))
```


###  Regression with Transformations

Sometimes, variable transformations (e.g. `log()` or `sqrt()`) are useful when we're doing linear regression, in order to get a better fit

```{r}
ggplot(Cars93) + aes(x = Fuel.tank.capacity, y = log(Price)) + 
  geom_point() + geom_smooth(method = lm, se = FALSE)
```


###  SEs with Transformations
I added `se = FALSE` so that the standard error bars do not show up in the graph.  Setting this parameter to `TRUE` instead, we get:

```{r}
ggplot(Cars93) + aes(x = Fuel.tank.capacity, y = log(Price)) + 
  geom_point() + geom_smooth(method = lm, se = TRUE)
```


###  Adjusting SE Bars

We can specify different alpha-levels for our confidence bands in liner regression.  The default is 0.95.  What if we changed it to 0.9?  0.99?

```{r}
ggplot(Cars93) + aes(x = Fuel.tank.capacity, y = log(Price)) + 
  geom_point() + stat_smooth(method = lm, se = TRUE, level = 0.9)
ggplot(Cars93) + aes(x = Fuel.tank.capacity, y = log(Price)) + 
  geom_point() + stat_smooth(method = lm, se = TRUE, level = 0.99)
```


##  Non-Linear Trends

We can also do other types of modeling, e.g. local regression / loess smoothing

```{r}
ggplot(Cars93) + aes(x = Fuel.tank.capacity, y = log(Price)) + 
  geom_point() + geom_smooth(method = loess, se = TRUE)
```



##  Leave The Points.  Take The Regression Model.

We don't even need to plot the points to do this -- sometimes we might want to just plot the regression model:

```{r}
ggplot(Cars93) + aes(x = Fuel.tank.capacity, y = log(Price)) +
  geom_smooth(method = loess, se = TRUE)
```


As always, can adjust some parameters:

```{r}
ggplot(Cars93) + aes(x = Fuel.tank.capacity, y = log(Price)) +
  geom_smooth(method = loess, se = TRUE, fill = "blue", 
              colour = "darkblue", size = 2, alpha = 0.2)
```


#  Next time:

+  Pairs plots
+  Don't forget about marginal distributions!
+  2-D Kernel Density Estimation
+  Heat Maps


