---
title: "Lecture 25 R Demo -- So Much More"
author: "Sam Ventura"
date: "November 30, 2016"
output: 
  html_document:
    toc:  true
    toc_float:  true
    code_folding:  show
---


#  `R` Notebooks

This is one of the reasons why we updated `R` and RStudio this week.  `R` Notebooks are similar to Jupyter notebooks (Python).  

See here for more details:  [https://blog.rstudio.org/2016/10/05/r-notebooks/]

I HIGHLY encourage you to read the info at the above link.  There's a lot of really cool stuff you can do with these.


##  Python

Speaking of Python:  Did you know you can use `R` Markdown to run code in other languages?

```{python}
def fib(n):
  a, b = 0, 1
  for _ in xrange(n):
    yield a
    a, b = b, a + b
print(list(fib(11)))
```


##  SQL

You can also run SQL queries within your `R` Notebook:

(Note:  I'm not adding the ` `'s here, because I don't actually have a `SQL` database called `db` or `my_database`)


{SQL, connection = db}
SELECT * FROM my_database



##  bash
You can run bash scripts as well:

```{r, engine = 'bash'}
cd /Users/sam/Desktop/bike/
ls
```


##  PDF Documents

For those of you who like LaTeX, you can use RMarkdown to generate PDF documents in addition to HTML documents.  Just change:

`output: html_document` to `output: pdf_document`

Be careful though, because you need to have the appropriate LaTeX tools installed on your computer.  RMarkdown with HTML, on the other hand, will work on any computer.



#  `ggplot2` Extras

A new version of `ggplot2` was just released.  Here's a quick overview of some of the interesting features.

See here for more information:

+  [https://blog.rstudio.org/2016/11/14/ggplot2-2-2-0/]
+  [https://www.r-bloggers.com/announcing-ggforce-accelerating-ggplot2]


##  Subtitles and Captions

We can now add subtitles and captions to graphs.

```{r, warning = F, message = F}
library(ggplot2)
library(MASS)
data(Cars93)

ggplot(Cars93, aes(x = Fuel.tank.capacity, y = MPG.highway)) + 
  geom_point(aes(color = Type, size = Weight)) + 
  geom_smooth(se = F, method = "lm") + 
  labs(
    title = "Miles Per Gallon vs. Fuel Tank Capacity",
    subtitle = "Points colored by car type and sized by car weight",
    caption = "(Source:  Cars93 Dataset)",
    x = "Fuel Tank Capacity (Gallons)",
    y = "Highway Miles Per Gallon"
  )
```




##  Hexagonal Bin Plots

These can be used as an alternative to traditional density estimation via contour plots or heat maps.

```{r, warning = F, message = F}
#install.packages("hexbin")
library(tidyverse)
olive <- read_csv("https://raw.githubusercontent.com/sventura/315-code-and-datasets/master/data/olive_oil.csv")
ggplot(olive, aes(x = palmitic, y = palmitoleic)) + 
  geom_hex(bins = 20)

```


##  `ggforce` Package and `facet_zoom()`

`facet_zoom()` allows you to zoom in on specific parts of the plot.

```{r, warning = F, message = F}
#install.pacakges("ggforce")
library(ggforce)
shark <- read_csv("https://raw.githubusercontent.com/sventura/315-code-and-datasets/master/data/attacks_test.csv")

ggplot(shark, aes(y = Age, x = Year)) + 
  geom_point(aes(color = Type)) + 
  facet_zoom(x = Year %in% 2000:2016)

ggplot(Cars93, aes(x = Fuel.tank.capacity, y = MPG.highway)) + 
  geom_point(aes(color = Type, size = Weight)) + 
  geom_smooth(se = F, method = "lm") + 
  labs(
    title = "Miles Per Gallon vs. Fuel Tank Capacity",
    subtitle = "Points colored by car type and sized by car weight",
    caption = "(Source:  Cars93 Dataset)",
    x = "Fuel Tank Capacity (Gallons)",
    y = "Highway Miles Per Gallon"
  ) + facet_zoom(x = Type == "Compact")

```



#  Interactive Graphics Packages

##  Flex Dashboard

Flex Dashboard is a really nice way to integrate your RMarkdown konwledge with Shiny.  It creates great-looking dashboard-style outupt that can be used when creating interactive graphics.

I'm going to demo how to create a flex dashboard, and then show a few examples.

Here are some links on the `flexdashboard` package:

+  [https://blog.rstudio.org/2016/05/17/flexdashboard-easy-interactive-dashboards-for-r/]
+  [http://rmarkdown.rstudio.com/flexdashboard/]
+  [http://rmarkdown.rstudio.com/flexdashboard/examples.html]


##  Shiny Themes

A closely related concept to `flexdashboard` is "Shiny Themes" (`shinythemes`), which allow you to choose different themes for your Shiny apps.

Here are some links on the `shinythemes` package:

+  [https://blog.rstudio.org/2016/10/13/shinythemes-1-1-1/]
+  [https://rstudio.github.io/shinythemes/]



##  Plotly

Plotly is **the best** package to make interactive graphics in `R` both within and outside of `shiny`.  It uses D3.js on the backend, so there's a lot of syntax that you may be unfamiliar with when comparing to `ggplot2` and some of the other `R` packages we've used this year.  Luckily, there's a free online book with a TON of example code (see links below) that should make it fairly easy to integrate into your final project.

You can easily use the `ggplotly()` function to create interactive versions of static graphics that have some features automatically included (e.g. alt-text).

You can also use the `plot_ly()` function to create new interactive graphs.

First, some links:

+  This is a comprehensive look into what you can do with Plotly in the form of a (free) book, including a ton of example code:  [https://cpsievert.github.io/plotly_book/index.html]
+  Here's the GitHub page for `plotly`, including a vignette with many examples:  [https://github.com/ropensci/plotly]


One nice feature of `plotly` is "linked brushing", which is often also referred to as "ribbons" or "linked views" or "multiple views" or "polygons".  I'll do a demo of this from the book above, and another demo using data from the [2016 Presidential Election](http://rpubs.com/cpsievert/229092).



##  HTML Widgets

HTML Widgets are a superset of `plotly`. They are in-development additions to the ever-growing list of possibily ways to make awesome charts in `R`.

Because these are being developed, updated, and improved, I don't know how to use all of these.  By the end of your final project, you will certainly know more about some of these topics than I will.

Some links:

+  HTML Widgets Showcase:  [http://www.htmlwidgets.org/showcase_metricsgraphics.html]
+  HTML Widgets Gallery:  [http://gallery.htmlwidgets.org/]




##  Custom JavaScript Widgets in Shiny

Create custom JavaScript widgets for Shiny!  Watch the tutorials described here:  [https://blog.rstudio.org/2016/05/06/shiny-javascript-tutorials/]


##  Miscellaneous

+  Here's a post about how you can use the `ggiraph` pacakge to make interactive graphics relating to the 2016 Olympics:  [http://blog.revolutionanalytics.com/2016/08/interactive-illustrator-quality-graphics-with-r.html]
+  Here's more info on that package:  [https://mran.microsoft.com/package/ggiraph/]
+  Here's the code for that analysis:  [https://github.com/d-qn/2016_08_02_rioOlympicsAthletes]




