---
title: "Lecture 23 R Demo -- Text"
author: "Sam Ventura"
date: "November 21, 2016"
output: 
  html_document:
    toc:  true
    toc_float:  true
    code_folding:  show
---


#  Introduction to Text Data

The graphs below don't have proper titles, axis labels, legends, etc.  Please take care to do this on your own graphs.

***

#  Terminology

**Corpus** = a collection of written texts, typically stored electronically in some structured way for processing.

+  Plural:  corpora
+  Etymology:  Latin "Corpor" = "body"; i.e. "body of texts"

**Documents** are individual texts in the corpus/collection.


***

**How do we summarize corpora?**  One common approach is using "document-term matrices":

*  Each row is an individual document in the corpus
*  Each column represents a unique word that occurs across all documents in the corpus
*  Cells can be filled in in several different ways, e.g.:
+  `X[i,j] = 1` if word $j$ occurs in document $i$, 0 otherwise
+  `X[i,j] = ` the "term frequency" of word $j$ in document $i$ (how many times does the word occur in the document?)
+  `X[i,j] = ` the TF-IDF score of word $j$ in document $i$

Note:  Some folks transpose the document-term matrix into a term-document matrix, where the rows are terms, and the columns are documents.

***

**TF-IDF = "term frequency / inverse document frequency"**:

+  Used to determine how "important" a word is to a document in a corpus.
+  Increases proportionally to the number of times a word appears in the document (the numerator)
+  Decreases proportionally to the number of times the word appears in the entire corpus (the denominator)
+  Adjusts for words that are generally more frequent

***

**Stop words**:  Extremely common words that are filtered out before text analysis (e.g. of, the, a, in, at, that, it, etc)

***

**Stemming**:  Reducing a word to a word stem, in order to capture the same word in different forms.  E.g.:

+  Reducing = reduc
+  Reduce = reduc
+  Reduced = reduc
+  Reducible = reduc
+  Reduction = reduc


***


**Sentiment**:  A view of or attitude toward a situation or event; an opinion.

+  "I love 315" = good sentiment
+  "Professor Ventura is the worst" = bad sentiment


***


**Sentiment Analysis**:  The process of computationally identifying and categorizing opinions expressed in a piece of text, especially in order to determine whether the writer's attitude towards a particular topic, product, etc., is positive, negative, or neutral.



***
***




#  Plotting Text

##  Most common words / phrases
##  Projections of the document-term matrix
##  Wordclouds
##  Sentiment Analysis

+  See the `tidytext` package and `tidytext::get_sentiments()`
+  See the "Tidy Text Mining in R" book for some examples!



#  Example Text Analysis -- Reuter's Mergers and Acquisitions Articles

You'll need the `tm` package.


```{r}
library(ggplot2)
library(tm)
data(acq)
#help(acq)

dtm <- DocumentTermMatrix(acq, control = list(weighting = weightTfIdf, 
                                              stopwords = TRUE))
class(dtm)
inspect(dtm[1:5,1:15])

n_mds <- 2
dtm2 <- as.data.frame((as.matrix(dtm)))
dtm_mds <- as.data.frame(cmdscale(dist(dtm2), n_mds))
names(dtm_mds) <- paste0("mds", 1:n_mds)
dtm_mds$index <- 1:nrow(dtm_mds)

ggplot(dtm_mds, aes(mds1, mds2)) + geom_text(aes(label = index))

acq[[9]]$content
```


***
***



#  Example Text Analysis -- Reuters Crude Oil Articles

You'll need the `tm` package.


```{r}
library(tm)
data(crude)
#help(crude)

dtm <- DocumentTermMatrix(crude, control = list(weighting = weightTfIdf, 
                                                stopwords = TRUE))
class(dtm)
inspect(dtm[1:5,1:15])

n_mds <- 2
dtm2 <- as.data.frame((as.matrix(dtm)))
dtm_mds <- as.data.frame(cmdscale(dist(dtm2), n_mds))
names(dtm_mds) <- paste0("mds", 1:n_mds)
dtm_mds$index <- 1:nrow(dtm_mds)

ggplot(dtm_mds, aes(mds1, mds2)) + geom_text(aes(label = index))

acq[[20]]$content
```




***
***



#  Example Text Analysis -- Reuters Crude Oil Articles vs. Mergers and Acquisitions Articles

You'll need the `tm` package.


```{r}
library(tm)
data(crude)
data(acq)
#help(crude)
#help(acq)


dtm_both <- DocumentTermMatrix(c(crude, acq), 
                               control = list(weighting = weightTfIdf, 
                                              stopwords = TRUE))

n_mds <- 20
dtm2 <- as.data.frame((as.matrix(dtm_both)))
dtm_mds <- as.data.frame(cmdscale(dist(dtm2), n_mds))
names(dtm_mds) <- paste0("mds", 1:n_mds)
dtm_mds$index <- c(paste0("C", 1:20), paste0("A", 1:50))
dtm_mds$type <- c(rep("C", 20), rep("A", 50))

#  Reorder the rows randomly
#  so that the "C" points aren't all covered by the "A" points
dtm_mds <- dtm_mds[sample(1:nrow(dtm_mds), nrow(dtm_mds)),]

ggplot(dtm_mds, aes(mds19, mds20)) + geom_text(aes(label = index, color = type))
```


***
***

#  Word Clouds

You'll need the `wordcloud` package.


#  Because I Couldn't Resist:  315 Mid Semester Feedback



```{r}
library(gsheet)
library(wordcloud)
feedback <- as.data.frame(gsheet2tbl("https://docs.google.com/spreadsheets/d/17tJCaT-2gyd_bAX1q5MPy5rCVqFzWTksVkD9_l-n4hs/edit?usp=sharing"))

#  Convert the vector in the data.frame into a corpus
feedback2 <- Corpus(VectorSource(as.character(feedback$text))) 

#  Convert to a plain text document
feedback_plain <- tm_map(feedback2, PlainTextDocument)

#  Remove punctuation
feedback_plain <- tm_map(feedback_plain, removePunctuation)

#  Remove stop words
feedback_plain <- tm_map(feedback_plain, removeWords, stopwords('english'))

#  Create a word cloud!
wordcloud(feedback_plain, max.words = 100, random.order = FALSE)


############################################################################

#  Any group structure?
dtm_315 <- DocumentTermMatrix(feedback2, 
                               control = list(weighting = weightTfIdf, 
                                              stopwords = TRUE))

n_mds <- 20
dtm2 <- as.data.frame((as.matrix(dtm_315)))
dtm_mds <- as.data.frame(cmdscale(dist(dtm2), n_mds))
names(dtm_mds) <- paste0("mds", 1:n_mds)
dtm_mds$index <- 1:nrow(dtm_mds)

ggplot(dtm_mds, aes(mds9, mds11)) + geom_text(aes(label = index))
feedback$text[98]
```


