---
title: "36-315 Lab 01 Solutions"
author: "Solutions"
date: "Due Friday, September 2, 2016 (6:30pm)"
output: html_document
---


##  Lab 01:  Hello, world!  Introduction to R, RStudio, & RMarkdown

Style Guide: Google's R Style Guide

###  Problem 1

```{r}
print("Hello, World!")
# I am printing 'Hello, World!' in R Markdown.
```

###  Problem 2

```{r}
#a. 
(6+3)*4-5

#b. 
4^2

#c.
exp(-5)

#d.
y <- 8
x <- 5-6*y
x

#e.
y <- 0
x <- 5-6*y
x

y <- 1
x <- 5-6*y
x

y <- 5/6
x <- 5-6*y
x
```


###  Problem 3

a.  By using help(quantile), the description of the ```probs``` argument is "numeric vector of probabilities with values in [0,1]. (Values up to 2e-14 outside that range are accepted and moved to the nearby endpoint.)".  

b.  By using help(mean) in the command line, the code at the bottom of the help page for the 'mean' function is:   

```{r, eval = FALSE} 
x <- c(0:10, 50)  
xm <- mean(x)  
c(xm, mean(x, trim = 0.10))
```

c. The name of the function to find the standard deviation is ```sd```.


###  Problem 4

a.  
```{r}
library(MASS)
library(help = MASS)
```
The official name of the ```MASS``` package is "Support Functions and Datasets for Venables and Ripley's MASS".  

b.  
```{r}
library(datasets)
library(help = "datasets")
```
The ```trees``` dataset has information about the girth, height and volume for black cherry trees.  

c.  
```{r}
library(graphics)
library(help = "graphics")
```

The function that creates a 1-D scatterplot is ```stripchart``` and using the ```help(stripchart)``` command, the argument ```x``` is "the data from which the plots are to be produced".


###  Problem 5

a.  
```{r}
lab1data <- read.table("https://raw.githubusercontent.com/sventura/315-code-and-datasets/master/data/lab1.txt", header = T)
head(lab1data)
```  

The function ```head``` displays the first six rows of the dataset.

b.  
```{r}
nrow(lab1data)
```  

```lab1data``` has 27 rows.

c.  
```{r}
lab1data <- read.table("https://raw.githubusercontent.com/sventura/315-code-and-datasets/master/data/lab1.txt", header = T, nrows = 10)
```

Adding ```nrow = 10``` allows you to specify the number of rows to read from the dataset.

d.  
```{r}
ncol(lab1data)
```

```lab1data``` has 10 columns.

e.  
```{r}
dim(lab1data)
```

The ```dim``` function returns both the number of rows and columns.

###  Problem 6

a.  
```{r}
myvector <- c(1,4,9,16,25,36,49,64,81,100,121,144,169,196)
```

b.
```{r}
row4 <- lab1data[4,]
min(row4)
max(row4)
```


c.  
```{r}
column4 <- lab1data[,4]
min(column4)
max(column4)
```


d.  
```{r}
lab1data[5,3]
lab1data[c(1,2),c(3,4)]
```


e.  
```{r}
col1 <- lab1data[,1]
mean(column4[which(col1 == 1)])
min(col1[which(column4 < 100)])
```

